Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans

### 2020-05-31 paper_PLOS/paper_PLOS_draft.tex
Added Figure, fixed figures 0-5

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{pdflscape}

@@

)

@@

\begin{longtable}[]{@{}lrlllr@{}}
\caption{Estimated Cancer Succeptibilty for nodes in
Atlantogenata}\tabularnewline
\toprule
Node & Est. Lifespan & K1 & K2 & Change in K & log2
Change\tabularnewline
\midrule
\endfirsthead
\toprule
Node & Est. Lifespan & K1 & K2 & Change in K & log2
Change\tabularnewline
\midrule
\endhead
Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
2.31\tabularnewline
Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 &
4.07\tabularnewline
Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
-2.31\tabularnewline
Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
0.00\tabularnewline
Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 &
27.66\tabularnewline
Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 &
2.31\tabularnewline
Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 &
4.14\tabularnewline
Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\tabularnewline
Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 &
-2.31\tabularnewline
Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 &
0.00\tabularnewline
Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 &
-23.05\tabularnewline
Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 &
29.97\tabularnewline
Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 &
6.92\tabularnewline
Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 &
8.62\tabularnewline
Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 &
0.00\tabularnewline
Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 &
-5.01\tabularnewline
Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 &
9.22\tabularnewline
Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 &
5.38\tabularnewline
Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
-4.61\tabularnewline
Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 &
-4.61\tabularnewline
Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 &
0.00\tabularnewline
Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 &
5.22\tabularnewline
Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 &
9.22\tabularnewline
Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 &
20.75\tabularnewline
Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 &
-3.76\tabularnewline
Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 &
1.51\tabularnewline
Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 &
0.00\tabularnewline
Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 &
0.00\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
\subsection*{Identification and evolutionary history of gene
duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}
\addcontentsline{toc}{subsection}{Identification and evolutionary
history of gene duplications}

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\includegraphics[width=6in]{paper_PLOS_draft_files/figure-latex/Figure0-1} \hfill{}

\caption{Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]}\label{fig:Figure0}
\end{figure}

@@

\begin{table}

\caption{\label{tab:Fig2 Prep Table 1}Estimated Cancer Succeptibilty for nodes in Atlantogenata}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrlllr}
\toprule
Node & Est. Lifespan & K1 & K2 & Change in K & log2 Change\\
\midrule
Loxodontini & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
Loxodonta africana & 65.00 & 2.47e+17 & 1.47e+16 & 1.68e+01 & 4.07\\
Loxodona & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
Loxodonta cyclotis & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
Palaeoloxodon antiquus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
\addlinespace
Elephantidae & 31.12 & 2.97e+15 & 1.40e+07 & 2.13e+08 & 27.66\\
Elephantina & 34.38 & 1.47e+16 & 2.97e+15 & 4.94e+00 & 2.31\\
Elephas maximus & 65.50 & 2.58e+17 & 1.47e+16 & 1.76e+01 & 4.14\\
Mammuthus & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
Mammuthus primigenius & 31.12 & 2.97e+15 & 1.47e+16 & 2.02e-01 & -2.31\\
\addlinespace
Mammuthus columbi & 34.38 & 1.47e+16 & 1.47e+16 & 1.00e+00 & 0.00\\
Proboscidea & 9.41 & 1.40e+07 & 1.21e+14 & 1.15e-07 & -23.05\\
Mammut americanum & 34.38 & 1.47e+16 & 1.40e+07 & 1.05e+09 & 29.97\\
Tethytheria & 25.49 & 1.21e+14 & 1.01e+12 & 1.21e+02 & 6.92\\
Trichechus manatus & 69.00 & 4.77e+16 & 1.21e+14 & 3.93e+02 & 8.62\\
\addlinespace
Paenungulata & 18.91 & 1.01e+12 & 1.01e+12 & 1.00e+00 & 0.00\\
Procavia capensis & 14.80 & 3.13e+10 & 1.01e+12 & 3.11e-02 & -5.01\\
Pseudoungulata & 18.91 & 1.01e+12 & 1.69e+09 & 5.97e+02 & 9.22\\
Orycteropus afer & 29.80 & 4.19e+13 & 1.01e+12 & 4.17e+01 & 5.38\\
Elephantulus edwardii & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
\addlinespace
Afrosoricida & 10.40 & 6.90e+07 & 1.69e+09 & 4.09e-02 & -4.61\\
Chrysochloris asiatica & 10.40 & 6.90e+07 & 6.90e+07 & 1.00e+00 & 0.00\\
Echinops telfairi & 19.00 & 2.57e+09 & 6.90e+07 & 3.72e+01 & 5.22\\
Afrotheria & 12.69 & 1.69e+09 & 2.83e+06 & 5.97e+02 & 9.22\\
Xenarthra & 20.89 & 4.97e+12 & 2.83e+06 & 1.76e+06 & 20.75\\
\addlinespace
Dasypus novemcinctus & 22.30 & 3.67e+11 & 4.97e+12 & 7.37e-02 & -3.76\\
Choloepus hoffmanni & 41.00 & 1.42e+13 & 4.97e+12 & 2.85e+00 & 1.51\\
Atlantogenata & 8.52 & 2.83e+06 & 2.83e+06 & 1.00e+00 & 0.00\\
Afroinsectivora & 12.69 & 1.69e+09 & 1.69e+09 & 1.00e+00 & 0.00\\
\bottomrule
\end{tabular}}
\end{table}

@@

\hypertarget{identification-and-evolutionary-history-of-gene-duplications}{%
\subsection{Identification and evolutionary history of gene
duplications}\label{identification-and-evolutionary-history-of-gene-duplications}}

\begin{table}

\caption{\label{tab:Figure3 Prep Genome Table}Summary of duplications in Atlantogenata}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllrrllr}
\toprule
Species & Common Name & Size (g) & \#Hits & \#Duplicated & \% Genes Found & \% Hits Duplicated & Mean ECNC/Hit\\
\midrule
Choloepus hoffmanni & Hoffmans Two-Toed Sloth & 4.3e+03 & 14082 & 3204 & 78.19\% & 22.75\% & 0.98\\
Chrysochloris asiatica & Cape Golden Mole & 49 & 13547 & 2716 & 75.22\% & 20.05\% & 0.99\\
Dasypus novemcinctus & Nine-Banded Armadillo & 4.8e+03 & 13819 & 2605 & 76.73\% & 18.85\% & 0.98\\
Echinops telfairi & Lesser Hedgehog Tenrec & 87 & 12903 & 1670 & 71.64\% & 12.94\% & 0.99\\
Elephantulus edwardii & Cape Elephant Shrew & 49 & 12884 & 3048 & 71.53\% & 23.66\% & 0.99\\
\addlinespace
Elephas maximus & Asian Elephant & 3.3e+06 & 14073 & 907 & 78.14\% & 6.44\% & 1.00\\
Loxodonta africana & African Savanna Elephant & 4.6e+06 & 14051 & 940 & 78.01\% & 6.69\% & 1.00\\
Loxodonta cyclotis & African Forest Elephant & 4.7e+06 & 14065 & 900 & 78.09\% & 6.40\% & 1.00\\
Mammut americanum & American Mastodon & 6e+06 & 13840 & 737 & 76.84\% & 5.33\% & 1.00\\
Mammuthus columbi & Columbian Mammoth & 6.6e+06 & 13059 & 426 & 72.51\% & 3.26\% & 1.00\\
\addlinespace
Mammuthus primigenius & Woolly Mammoth & 4.3e+06 & 13935 & 723 & 77.37\% & 5.19\% & 1.00\\
Orycteropus afer & Aardvark & 5.3e+04 & 13880 & 1083 & 77.06\% & 7.80\% & 0.99\\
Palaeoloxodon antiquus & Straight Tusked Elephant & 1e+07 & 13969 & 745 & 77.56\% & 5.33\% & 1.00\\
Procavia capensis & Rock Hyrax & 3e+03 & 13672 & 788 & 75.91\% & 5.76\% & 1.00\\
Trichechus manatus & Manatee & 6.9e+05 & 14092 & 1046 & 78.24\% & 7.42\% & 1.00\\
\bottomrule
\end{tabular}}
\end{table}

\includegraphics[width=6in,height=7in]{paper_PLOS_draft_files/figure-latex/Figure3-1} \hfill{}

\caption{Gene duplications occur readily throughout Atlantogenata. A) A table summarizing duplication events in Atlantogenata. B) A tree of the species in Atlantogenata with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.}\label{fig:Figure3}
\end{figure}

Enhanced cancer suppression may have evolved through many mechanisms;
among the most parsimonious is an increase in the copy number of genes
with tumor suppressor functions. Previous studies focusing on candidate
gene studies, for example, have identified increased copy number of the
tumor suppressors \emph{TP53} and \emph{LIF} in elephants
{[}12,21--24{]}. Therefore, in order to test whether this was pervasive
genome-wide in \emph{Afrotherians}, we used a Reciprocal Best Hit BLAT
(RBHB) approach to infer gene copy number in \emph{Afrotherian} and
\emph{Atlantogenatan} genomes (Fig. 2A, Supplementary Figure 1). Because
RBHB-like approaches can over-estimate copy number when genes are
fragmented or incorrectly assembled across multiple scaffolds, we also
inferred copy number using a complementary method that quantifies the
ratio between observed and expected gene coverage per nucleotide (ECNC)
(Supplementary Figure 1). By only including nucleotides from the query
sequence that were observed in the target genome, we also correct for
partial hits where some or all of the homologs of a gene have diverged
from the human homolog.

As our sequence database included various protein transcripts for each
gene, in order to obtain gene-level copy number information and
eliminate any many-to-one mappings of hits, we labeled each exon of
every reciprocal best hit (RBH) with the gene corresponding to the query
transcript and merged all overlapping exons; next, we eliminated any
many-to-one exons that resulted from the previous step. Finally, we
reassembled the gene loci based on the original transcript starts and
ends, and the collapsed exon data, obtaining the full sequence of each
RBH locus. Genes were considered to be duplicated if its copy number via
RBHB was greater than or equal to 2, and the maximum ECNC among all
transcripts prior to filtering was greater than or equal to 1.50 in
order to account for partial hits that may be due to incomplete
coverage.

Once we applied these criteria to the results, we obtained the results
shown in Figure 2A. Our approach positively identified an average of NA
of the queries we searched. We observed that the percentage of
duplicated genes in non-Pseudoungulatan genomes was significantly
higher: while these genomes had duplicates percentages ranging anywhere
from 3.26\% to 7.80\%, outgroup species' duplication rates ranged from
12.94\% to 23.66\%.

{\centering \includegraphics[width=5in,height=5in]{paper_PLOS_draft_files/figure-latex/Figure4-1}

}

\caption{Correlations between genome quality metrics and ECNC metrics}\label{fig:Figure4}
\end{figure}

Given the preliminary status of the genomes for \emph{Chrysochloris
asiatica} and \emph{Elephantulus edwardii}; and the fact that more
revisions have been performed for Paenungulatan genomes; we hypothesized
that genome quality may be confounding the results we've observed. We
used \texttt{CEGMA} and the \texttt{gVolante} server {[}31,32{]} to
assess the quality of all our genomes, to see if any of these metrics
correlated with our metrics of gene counts and copy number. As shown in
Figure 3, mean ECNC, mean Copy Number, and mean CN (the lesser of Copy
Number and ECNC per gene) correlate moderately strongly with genomic
quality metrics related to the length and assembly quality, like LD50
and the number of scaffolds and contigs with a length above either 100K
or 1M, supporting our hypothesis that increases in copy number correlate
.

In order to determine when in the phylogeny our identified duplications
occurred, we used maximum likelihood to reconstruct ancestral gene copy
numbers across the phylogeny. We encoded the copy number of each gene as
a discrete trait ranging from 0 (``?'') to 31+, and then ran the
ancestral gene copy number reconstruction using the phylogenetics suite
\texttt{IQTREE}, which rapidly tests and runs the best evolutionary
model for a given trait of interest. Importantly, \texttt{IQTREE} can be
set to only output states where the likelihood of the state is greater
than a certain threshold. Setting the threshold to 0.8, we obtained the
results shown in Figure 2B.

Gene duplications events occured readily thoughout \emph{Atlantogenata}
(Figure 2B). Among the genes that increased in copy number in the
elephant lineage are TP53 and LIF, as previously described; however, we
find that these two genes represent a fraction of the 940 genes that are
duplicated in the African Elephant overall, which accumulated over
various steps through their evolution. While the extinct elephantids
have acceptable genome quality metrics according to CEGMA, they are
nonetheless missing a significant number of sequences; this may
contribute to the low number of duplicated genes that occured in
internal nodes. The number of duplicates that occur at each branch is
also proportional to the density of the sampling of the clade overall,
as would be expected. In branches, such as \emph{Afrosoricida}, where
the number of species is relatively minuscule compared to the size of
the clade, we see many significantly larger numbers of duplications
private to these species.

\hypertarget{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}{%
\subsection{Duplications that occurred recently in Probodiscea are
enriched for tumor suppressor
pathways}\label{duplications-that-occurred-recently-in-probodiscea-are-enriched-for-tumor-suppressor-pathways}}

\begin{table}

\caption{\label{tab:Branches which experienced an increase in size}Number of pathways overrepresented among duplicated genes at different FDRs.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrrrr}
\toprule
Ancestor & Node & Pathways at FDR<0.1 & Pathways at FDR<0.2 & Pathways at FDR<0.3 & Pathways at FDR<0.5\\
\midrule
Afroinsectivora & Elephantulus edwardii & 252 & 37 & 30 & 87\\
Afrosoricida & Chrysochloris asiatica & 90 & 48 & 43 & 105\\
Afrosoricida & Echinops telfairi & 0 & 2 & 0 & 31\\
Afrotheria & Afroinsectivora & 0 & 0 & 0 & 33\\
Loxodontini & Loxodonta africana & 6 & 0 & 1 & 0\\
\addlinespace
Paenungulata & Tethytheria & 0 & 0 & 0 & 2\\
Proboscidea & Mammut americanum & 0 & 0 & 0 & 6\\
Pseudoungulata & Orycteropus afer & 27 & 67 & 29 & 67\\
Tethytheria & Proboscidea & 0 & 3 & 0 & 3\\
Tethytheria & Trichechus manatus & 4 & 0 & 0 & 2\\
\bottomrule
\end{tabular}}
\end{table}

\includegraphics[width=6in,height=4in]{paper_PLOS_draft_files/figure-latex/Figure5-1} \hfill{}

\caption{Overrepresentation Analysis of Duplicated Genes in Atlantogenata using Reactome Pathways.}\label{fig:Figure5}
\end{figure}

Our initial hypothesis was that genes which duplicated in lineages that
experienced a growth in size would be enriched for membership in tumor
suppressor pathways. Thus, we used WebGestalt and its Overrepresentation
Analysis (ORA) functionality to determine what pathways were enriched in
our duplicated gene sets in each branch relative to our initial query
set. For our database, we used Reactome for our primary analysis, but
additionally used the KEGG, Panther, Wikipathways, and
Wikipathways\_cancer databases using WebGestalt ORA. Going through the
tree, at no FDR\textless{}0.5 is there any significant pathway
representation for genes that increased in the branches leading to , or
\emph{Paenungulata}; furthermore, there are no significant pathway
enrichments at FDR\textless{}0.5 for genes whose copy number did not
change between branches (copy-number-stable) for , or \emph{Afrotheria}.
Note that because \emph{Xenarthra} was selected as the outgroup, it is
not possible to polarize the changes in their gene copy numbers along
the tree.

For the other branches, the number of pathways that came up as
significantly enriched at each FDR is shown in Figure 3A (Supplementary
Figure 3). For the species with high duplication rates and
lower-quality, highly-fragmented genomes, such as with
\emph{Chrysochloris asiatica} (20.05\% duplicated hits) and
\emph{Elephantulus edwardii} (23.66\% duplicated hits), it is
unsurprising that there is a proportionally large number of pathway
enrichments. In the case of these two species, their many pathway
enrichments also span an incredible range of processes at every level of
biology; this, in combination with the high number of copy numbers
identified for these genes, further suggests a need for improvement and
refinement in these genomes. In the cell cycle pathways called as
significant in the genomes of these two species plus \emph{Orycteropus
afer} and \emph{Echinops telfairi}, the duplicated genes included in
these sets are from the same gene families, such as the APC subunit
family; the proteosome subunit families; and the protein phosphatase 2
family, among others. It is highly possible that these results reflect
true expansions of these gene families, especially in the higher-quality
\emph{Orycteropus afer} genome; however, it is also possible that it
simply reflects artificial duplication events, and so require further
study.

The pathway enrichments for genes whose copy number either did not
change, or whose copy number increased, between \emph{Loxodona} and the
African Elephant are shown in Figure 3B and 3C, respectively. Among the
few enriched pathways in the African Elephant, we see that that two
tumor suppression pathways - APC Complex-related pathways, and ``TP53
Regulates Metabolic Genes''- appear not only in the case of stable
genes, but also in the set of newly duplicated genes. The other pathways
we see in the set of recently-duplicated genes include
``Functionalization of Compounds'' and its daughter pathway
``Xenobiotics''. Genes in these pathways serve to add functional groups
to lipophylic compounds which would otherwise not be reactive in the
cell, and are types of metabolic pathways. In the stable set we see
enrichment of pathways such as ``Neuronal Systems'' and ``Axon
Guidance,'' which fit in well with what is known about elephant biology
and evolution {[}47{]}. Overall in elephants, we see enrichments within
duplicated genes for pathways involved in what makes an elephant an
elephant - including tumor suppressor pathways.

\hypertarget{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}{%
\subsection*{\texorpdfstring{Concerted duplication of TP53 and
TP53-related genes towards
\emph{Probodiscea}}{Concerted duplication of TP53 and TP53-related genes towards Probodiscea}}\label{concerted-duplication-of-tp53-and-tp53-related-genes-towards-probodiscea}}
\addcontentsline{toc}{subsection}{Concerted duplication of TP53 and
TP53-related genes towards \emph{Probodiscea}}

@@

\leavevmode\hypertarget{ref-Goodman2009}{}%
47. Goodman M, Sterner KN, Islam M, Uddin M, Sherwood CC, Hof PR, et al.
Phylogenomic analyses reveal convergent patterns of adaptive evolution
in elephant and human ancestries. Proceedings of the National Academy of
Sciences. 2009;106: 20824--20829.
doi:\href{https://doi.org/10.1073/pnas.0911239106}{10.1073/pnas.0911239106}
